Follow the Flow of Rain and Paint: An Instance Optimal Contour Tree Algorithm

### 2014-06-10 paint-contour-tree.tex
writing paint section

\newcommand{\cut}{{\tt cut}}
\newcommand{\cont}{\psi}
\newcommand{\lift}{{\tt lift}}
\newcommand{\paint}{{\tt paint}}
\newcommand{\rain}{{\tt rain}}
\newcommand{\reeb}{R}
\newcommand{\surgery}{{\tt surgery}}
\newcommand{\wet}{{\tt wet}}

@@

We begin by associating a color with each maximum. If the set of maxima is $S$,
this is simply an arbitrary bijection $c:S \mapsto [|S|]$. Imagine there being a large
can of paint of color $c(y)$ at maxima $y$. We will spill different paint from each maximum and watch it flow down.
This is analogous to the raining in the previous section, but paint is a much more viscous liquid.
\emph{So paint only flows down edges, and it does not color the interior of faces.} Furthermore, paints
do not mix, so eventually, every vertex and edge of $\MM$ gets a unique color.

We begin with the procedure \paint. We can think of this procedure as extending the map $c:S \mapsto [|S|]$
to $c:G(\MM) \mapsto [|S|]$.

>>>>

\newcommand{\cut}{{\tt cut}}
\newcommand{\cont}{\psi}
\newcommand{\lift}{{\tt lift}}
\newcommand{\paint}{{\tt paint}}
\newcommand{\rain}{{\tt rain}}
\newcommand{\reeb}{R}
\newcommand{\surgery}{{\tt surgery}}
\newcommand{\touch}{T}
\newcommand{\wet}{{\tt wet}}

@@

We begin by associating a color with each maximum. If the set of maxima is $S$,
this is simply an arbitrary bijection $\chi:S \mapsto [|S|]$. Imagine there being a large
can of paint of color $c(y)$ at maxima $y$. We will spill different paint from each maximum and watch it flow down.
This is analogous to the raining in the previous section, but paint is a much more viscous liquid.
\emph{So paint only flows down edges, and it does not color the interior of faces.} Furthermore, paints
do not mix, so eventually, every vertex and edge of $\MM$ gets a unique color.
%
\begin{definition} \label{def:paint} A  \emph{painting} of $\MM$ is a map $\chi:V \cup E \mapsto [|S|]$
with the following properties. (Remember that $G(\MM) = (V,E)$.)
\begin{asparaitem}
	\item The restriction $c:S \mapsto [|S|]$ is a bijection.
	\item Consider an edge $e$. There exists a descending path from some maximum $x$ to $e$
	consisting of edges in $E$, such that all edges and vertices along this path have the same color. (The maximum $x$
	is well-defined to be the unique maximum with the same color as $e$.)
\end{asparaitem}
\end{definition}
%
Note that paintings are not unique. It is straightforward to construct a painting of $\MM$ in $O(n)$ time, using a descending BFS from
each maximum. 
%
\begin{definition} \label{def:color-set} Fix a painting $\chi$, and a color $c$.
A critical point $x$ is \emph{$c$-touched} if $x$ is incident to a $c$-colored
increasing edge. The set of $c$-touched critical points is denoted $\touch(c)$.
\end{definition}
%
We will store each set $\touch(c)$ in a binomial heap. Abusing notation, $\touch(c)$ refers
both to the set and the data structure used to store it.
